Homework Assignment 2 \\ Loss Functions and Support Vector Machines

### 2017-02-27 homeworks/hw2.tex
Homework 2

\title{Homework Assignment 1 \\
Perceptron and Logistic Regression}
\author{Lecturer: Kyunghyun Cho}
%\affil{
%    Courant Institute of Mathematical Sciences and \\
%    Center for Data Science,\\
%    New York University 
%}

@@

\paragraph{Submission Instruction}

You must typeset the answers to the theory questions using LaTeX or Microsoft
Word and compile them into a single PDF file. For the programming assignment,
complete the two Jupyter notebooks.  Create a
ZIP file containing both the PDF file and the two completed Jupyter notebooks,
and name it ``$\left<\text{Your-NetID}\right>$\_hw1.zip''. Email this file to
\url{intro2ml@gmail.com} within two weeks since the announcement of the
homework.

\paragraph{1.} When defining a perceptron, we have augmented an input vector
$\vx$ with an extra $1$: 
\begin{align*}
    M(\vx) = \sign(\vw^\top \tilde{\vx}),
\end{align*}
where $\tilde{\vx} = \left[ \vx; 1\right]$. Why is this necessary? Provide an
example in which this extra $1$ is necessary.
\\
\\
SOLUTION: The model equation created after the learning process is supposed to provide a line that classifies inputs as accurately as possible. The number of variables is equivalent to $d$, or the dimension of $\vx$, and each coefficient corresponds to the values in $\vw$. With such an equation, there is nowhere to place a constant to account for a vertical shift in the model: it only permits models which pass through the origin. In some cases, the optimal solution after learning may be a model that does not pass through the origin, but somewhere else along the y axis. By appending a 1 to the end of $\vx$, we can obtain an additional value from $\vw$ during learning, which ends up being the bias. This "bias" helps shift the model by a constant and can thus provide a more optimal solution.

@@

\paragraph{2.} We used the following distance function for perceptron in the
lecture:
\begin{align*}
    D(M^*(\vx), M, \vx) = -\left( M^*(\vx) - M(\vx)
    \right) \left(\vw^\top
    \tilde{\vx}\right).
\end{align*}
This distance function has a problem of a trivial solution. What is the trivial
solution? Propose a solution to this.
\\
\\
SOLUTION: The trivial solution is when we do not want to assign different weights to the input vector. As an example, if W is 0, then that means the distance function will always output 0, even if it shouldn't. To solve this, we need to somehow augment the weight vector to account for this situation and add a sort of intercept.

@@

\paragraph{3.} The distance function of logistic regression was defined as 
\begin{align*}
    D(y^*, \vw, \vx) = -(y^* \log M(\vx) + (1-y^*) \log (1- M(\vx))).
\end{align*}
Derive its gradient with respect to the weight vector $\vw$ step-by-step.
\\
\\
SOLUTION: 
Let $a = -y^* \log M(\vx)$ and $b =  \log (1- M(\vx))$ \\
We also know that $M(\vx) = \sigmoid(\vw^\top \tilde{\vx}) $ \\
Based off this, we can say $a = -y^* \log ( \frac{1}{1+e^{-\vw^{\top \tilde{\vx}}}})$ and $b =  \log (1- \frac{1}{1+e^{-\vw^\top \tilde{\vx}}})$ \\
$b = \log (1- \frac{1}{1+e^{-\vw^\top \tilde{\vx}}}) = \log ( \frac{1+e^{-\vw^\top \tilde{\vx} } - 1}{1+e^{-\vw^\top \tilde{\vx}}}) = \log ( \frac{e^{-\vw^\top \tilde{\vx} }}{1+e^{-\vw^\top \tilde{\vx}}}) =  \log ( e^{-\vw^\top \tilde{\vx} } ) - \log (1+e^{-\vw^\top \tilde{\vx}}) $ \\
$ \frac{db}{d \vw^\top} = \frac{-\tilde{\vx} e^{-\vw^\top \tilde{\vx}}}{e^{-\vw^\top \tilde{\vx}}} + \frac{\tilde{\vx} e^{-\vw^\top \tilde{\vx}}}{1 + e^{-\vw^\top \tilde{\vx}}}  = \frac{-\tilde{\vx} (1 + e^{-\vw^\top \tilde{\vx}}) +\tilde{\vx} e^{-\vw^\top \tilde{\vx}}}{1 + e^{-\vw^\top \tilde{\vx}}}  = \frac{\tilde{\vx}}{1 + e^{-\vw^\top \tilde{\vx}}}$ \\
$ \frac{da}{d \vw^\top} = \frac {-y^* \tilde{\vx}e^{-\vw^\top \tilde{\vx}}}{1 + e^{-\vw^\top \tilde{\vx}}}$ \\
Bringing this together: \\
$ \nabla_{\vw} D(M^*(\vx), M, \vx) =  \frac {-y^* \tilde{\vx}e^{-\vw^\top \tilde{\vx}}}{1 + e^{-\vw^\top \tilde{\vx}}} - \frac{y^* \tilde{\vx}}{1 + e^{-\vw^\top \tilde{\vx}}} + \frac{\tilde{\vx}}{1 + e^{-\vw^\top \tilde{\vx}}} $ \\
$ = -y^* \tilde{\vx}e^{-\vw^\top \tilde{\vx}} \sigmoid(\vw^\top \tilde{\vx}) -y^* \tilde{\vx}\sigmoid(\vw^\top \tilde{\vx}) + \tilde{\vx}\sigmoid(\vw^\top \tilde{\vx}) = \tilde{\vx}\sigmoid(\vw^\top \tilde{\vx}) (-y^* (1 + e^{-\vw^\top \tilde{\vx}}) + 1)$ \\
$ = \tilde{\vx}\sigmoid(\vw^\top \tilde{\vx}) (\frac{-y^* }{\sigmoid(\vw^\top \tilde{\vx})} + 1) = \tilde{\vx} (-y^* + \sigmoid(\vw^\top \tilde{\vx}))$ \\
$ = -(M^*(\vx) - M(\vx)) \tilde{\vx}$

\paragraph{4.} (Programming Assignment)
Complete the implementation of perceptron and logistic regression using Python
and scikit-learn. The completed notebooks must be submitted together with the
answers to the questions above. 
\begin{itemize}
    \item[Perceptron]
        \url{https://github.com/nyu-dl/Intro_to_ML_Lecture_Note/blob/master/notebook/Perceptron1.ipynb}
    \item[Logistic Regression]
        \url{https://github.com/nyu-dl/Intro_to_ML_Lecture_Note/blob/master/notebook/Logistic%20Regression%201.ipynb}
\end{itemize}

>>>>

\title{Homework Assignment 2 \\
    Loss Functions and Support Vector Machines
}
\author{Lecturer: Kyunghyun Cho}
%\affil{
%    Courant Institute of Mathematical Sciences and \\
%    Center for Data Science,\\
%    New York University 
%}

@@

\paragraph{1.} After replacing the label set from $\left\{ 0,1 \right\}$ to
$\left\{ -1,1 \right\}$, we introduced the log loss
\begin{align*}
    D_{\log}(y, \vx; M) = \frac{1}{\log 2} \log(1+\exp(-s(y, \vx; M))),
\end{align*}
as an alternative to the logistic regression distance function above. Show that
these two are equivalent up to a constant multiplication for logistic
regression.

@@

\paragraph{2.} Unlike the log loss, the hinge loss, defined below, is not
differentiable everywhere:
\begin{align*}
    D_{\hinge}(y, \vx; M) = \max(0, 1 - s(y, \vx; M)).
\end{align*}
Does it mean that we cannot use a gradient-based optimization algorithm for
finding a solution that minimizes the hinge loss? If not, what can we do about
it?
\\
\\
SOLUTION: For points $ > 1$, the derivative is 0. For points $< 1$, the hinge function is equal to $1 - y\vw^\top \tilde{\vx}$. This case has a trivial derivative of $-y \tilde{\vx}$. The only part which we cannot find the derivative is at the point where $s(y, \vx; M) = 1$. Instead, we can simply arbitrarily define the derivative to be 0 at that point because as far as we are concerned, a value should be considered correctly classified if it is equal to exactly 1. Anything greater is a correct classification and values slightly below are barely classified. 
\vspace{5mm}

\paragraph{3.} When the distances to the nearest positive and negative examples
are defined as $d^+$ and $d^-$, the margin is 
\begin{align*}
    \gamma =& \frac{1}{2} (d^+ + d^-).
\end{align*}
Show that minimizing the norm of the weight vector of a support vector machine
is equivalent to maximizing the margin.

@@

\paragraph{4.} \todo{PROGRAMMING ASSIGNMENT}
